---
documentclass: book
title:  Handbook for the dm R Package
author: "Robert Schnitman"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
site: bookdown::bookdown_site
output: 
  bookdown::gitbook: 
    config:
      toc:
        collapse: section
      search: yes
  bookdown::pdf_book: default
biblio-style: apalike
link-citations: yes
description: "The purpose of the dm R package is to provide data management tools as inspired by Stata, SPSS, and other statistical packages."
---

# Introduction {-#index}

**UNDER CONSTRUCTION**

The `dm` package is inspired by the data management functions in other statistical packages such as Stata and SPSS. The purpose of this package is to simplify data management/checking.

# Installation {-}

Currently, this package is only available on Github, so please use `devtools` to install this package.

```{r, eval = FALSE}
if (!require(devools)) {
  
  install.packages('devtools')
  library(devtools)
  
}

install_github('robertschnitman/dm')
```

```{r, message = FALSE}
library(dm)
```

# `between()`

The `between()` function acts similar to SQL's `BETWEEN` clause[^1], conditioning values between lower and upper limits.

The function has four parameters:

1. `x`, a vector;
1. `low`, a user-specified lower limit;
1. `high`, a user-specified upper limit; and
1. `inclusive = TRUE`, an input that determines whether the `between()` function will condition inclusive (if `TRUE`) of the specified limits or not (if `FALSE`).

A vector is returned by the function.

[^1]: https://www.w3schools.com/sql/sql_between.asp

```{r}
k <- 1:100
between(k, 5, 24)
```

```{r}
subset(mtcars, between(mpg, 18, 21, inclusive = FALSE))
```

# `dedup()`

The `dedup()` function deduplicates data. It functions the same as `unique()`.

```{r}
dedup(iris$Species)
```


## `dup_rows()`

The `dup_rows()` functions returns the duplicated rows in a 2-dimensional dataset. Functionally the same as `x[duplicated(x), ]`, where `x` is a matrix or data frame.

```{r}
dup_rows(iris)
```


## `dup_nrow()`

The `dup_nrow()` function counts the number of duplicates in a dataset.

```{r}
dup_nrow(iris)
```

## `dup_mark()`

The `dup_mark()` function codes duplicate rows with a 1 and non-duplicates with a 0.

```{r}
dup_mark(iris)
```


# `differ2()`

The `differ2()` function codes differences between two vectors as a 1; otherwise, 0. This function is based on Stata's `egen differ()` command. [^1]

[^1]: https://www.stata.com/manuals13/degen.pdf

```{r}
x <- 1:5
y <- c(1, 6, 3, 8, 5)

z <- differ2(x, y)

cbind(x, y, z)
```

# `expected()`

The `expected()` function stops an execution if a specific class is not expected.

```{r, error = TRUE}
expected(2, 'character')
```

# `keep()`

The `keep()` function behaves the same as `subset()`. The naming scheme is inspired by Stata's `keep` command.[^1]

[^1]: https://www.stata.com/manuals13/gsw12.pdf

```{r}
keep(mtcars, between(mpg, 18, 21))
```

# `load_libraries()`

The `load_libraries()` function checks whether a set of libraries have been installed: if the libraries exist, the function will load them; otherwise, the function will install and then load them.

```{r, eval = FALSE}
load_libraries('tidyverse', 'abind', 'ggformula')
```

# `numNA()`

Similar to Stata,[^1] numNA counts the number of missing values: if m is set to 1, then it counts the number of missing values per row; if set to 2, then it counts the number of missing values per column; otherwise, it counts the total number of missing values.

[^1]: https://www.stata.com/manuals13/m-5missing.pdf

```{r}
numNA(airquality)
```

## `rowNA()`/`colNA()`

The function `rowNA()` is the equivalent of `numNA(x, 1)`, where x is a data frame. It counts the number of missing values per row. The function `colNA()` does the same but per column.

```{r}
rowNA(airquality)
colNA(airquality)
```


# `parity()`

The function `parity()` determines whether a number is even or odd. This function and its related functions `is.even()` and `is.odd()` are inspired by Julia's `iseven()` and `isodd()` functions.[^1] The `get_even()` and `get_odd()` functions subset a vector for its even and odd numbers respectively.

[^1]: https://docs.julialang.org/en/release-0.4/stdlib/numbers/

```{r}
parity(mtcars$carb)
is.even(1)
is.even(mtcars$carb)
is.odd(2)
is.odd(mtcars$carb)
get_odd(mtcars$carb)
get_even(mtcars$carb)
```


# `recode()`

The `recode()` function replaces an initial set of values with a new set of values for a vector. The function `recode_all()` applies `recode()` to all columns in a dataset.

The inputs for `recode()` are the following:

1. a vector;
1. a set of values to replace ("initial values"); and
1. a set of values values that will replace the initial values.

```{r}
mtcars$am
recode(mtcars$am, 0:1, 2:3)
```

## `NAvl()`

The `NAvl()` function replaces a vector's missing values with a specified value set. The function `NAvl0` replaces a vector's missing values with 0.

```{r}
with(airquality, NAvl(Ozone, mean(Ozone, na.rm = TRUE)))
NAvl0(airquality$Ozone)
```

## `switchv()`

The `switchv()` function behaves the same as `switch()` except that it applies over a vector (i.e., it is a vectorized version of the latter). The function `swap` is a shorthand synonym for `switchv()`. The function `switchv_all()`/`swap_all()` applies `switchv()`/`swap()` to all columns in a data frame.

These functions are inspired by SPSS's `RECODE` command.[^1]

[^1]: https://libguides.library.kent.edu/SPSS/RecodeVariables

```{r}
switchv(iris$Species, setosa = 1, versicolor = 2, virginica = 3)
```

# `rename_file()`

The function `rename_file()` renames files in a directory based on a given pattern. It has three inputs:

1. `filepath`, which specifies the desired directory of where the files to be renamed are;
1. `pattern_now`, which is a regular expression pattern of the file names as they currently are; and
1. `pattern_new`, which is the desired string that replaces the file names detected in `pattern_now`.

```{r, eval = FALSE}
# From a folder, replace file names having "Fall 2018" with "Spring 2019."
rename_file('C:/my/folder/', 'Fall 2018', 'Spring 2019')
```


# `stopif()`

# `type_class()`

# `zero_flag()`

# References {-}

Julia documentation. *numbers*. https://docs.julialang.org/en/release-0.4/stdlib/numbers/

Kent State University. *SPSS Recode*. https://libguides.library.kent.edu/SPSS/RecodeVariables

Stata. *egen*. https://www.stata.com/manuals13/degen.pdf

w3schools.com. *SQL BETWEEN OPERATOR*. https://www.w3schools.com/sql/sql_between.asp

# See also {-}

`dm` GitHub Page. https://github.com/robertschnitman/dm

IBM's SPSS. https://www.ibm.com/analytics/spss-statistics-software

Robert Schnitman's Profile. https://robertschnitman.netlify.com/

Stata's website. https://www.stata.com/